PyTracer: Automatically Profiling Numerical Instabilities in Python

### 2022-05-16 IEEEtran/paper.tex

\ifANNOTATED
\DeclareRobustCommand{\add}[1]{\sethlcolor{lightgreen}\hl{#1}}
\DeclareRobustCommand{\remove}[1]{\textcolor{lightred}{\sout{#1}}}
\DeclareRobustCommand{\addCite}[1]{\textcolor{lightgreen}{\mbox{#1}}}
\newcommand{\TG}[1]{\color{red}\textsc{From Tristan}: #1\xspace\color{black}}
\else
\DeclareRobustCommand{\add}[1]{#1}
\DeclareRobustCommand{\remove}[1]{}
\DeclareRobustCommand{\addCite}[1]{#1}
\newcommand{\TG}[1]{}
\fi

@@

\pytracer estimates the variability of floating-point variables by combining
program traces obtained with different numerical perturbations, which requires
(\add{a}\remove{1}) generating numerically-perturbed executions,
(\add{b}\remove{2}) tracing floating-point computations along program
executions, and (\add{c}\remove{3}) visualizing stability evaluations.
\add{PyTracer addresses (a) by relying on the ``fuzzy" Python environment instrumented with Monte-Carlo
    arithmetic, (b) by dynamically instrumenting Python functions,
    modules, and classes through meta-programming, and (c) by visualizing stability measures throughout program executions
    with an interactive Plotly
    dashboard.
    \mbox{\pytracer} is available at
    \mbox{\href{https://github.com/big-data-lab-team/pytracer.git}{github.com/big-data-lab-team/pytracer.git}}
    under MIT License.} \remove{\mbox{\pytracer} addresses these requirements with
    (1) a ``fuzzy" Python interpreter instrumented with Monte-Carlo arithmetic, (2)
    a dynamic instrumentation of Python functions, modules, and classes through
    meta-programming, (3) an interactive Plotly dashboard to visualize stability
    measures throughout program executions. }

@@

% \add{PyTracer follows a three steps approach:}
% \begin{enumerate}
%     \item \add{Automatic instrumentation of Python programs to save traces containing input and output values of all functions.}
%     \item \add{Traces aggregation to produce a statistics summary.}
%     \item \add{Interactive visualization to navigate through code execution.}
% \end{enumerate}

\add{To detect numerical instabilities, \pytracer works in three main steps:}
\begin{enumerate}
\item It instruments Python programs and their module dependencies automatically, to save the input and output values of functions in a trace file.
\item It executes the instrumented program several times using Monte-Carlo Arithmetic to inject perturbations in function outputs.
\item It aggregates all trace files to compute statistics about the traced values, and visualize them in a GUI.
\end{enumerate}
\add{
    To implement these steps, \pytracer adopts the workflow shown in Figure~\ref{fig:workflow}.
At step 1, \pytracer uses \textit{monkey patching} to replace the functions called
by the application with a wrapper that saves their arguments and returned values
in a trace file. At step 2, it executes the application multiple times with Monte-Carlo Arithmetic as implemented in Fuzzy, a collection 
of scientific computing libraries recompiled with the Verificarlo tool. It then
aggregates the resulting trace files to compute summary statistics on
traced variable. At step 3, a Plotly dashboard provides interactive visualizations
highlighting unstable code sections.
}

% \add{
%     The main goal of PyTracer is to help developpers visualize the numerical
%     stability of Python applications. PyTracer saves time for the user by
%     instrumenting and aggregating values. In the same way that GDB makes the user
%     save time when debugging an application by providing all information in one
%     layout, PyTracer allows the user to visualize the overall numerical quality to
%     quickly detect instabilities. The PyTracer visualization layout improves result
%     analysis in any case since 1) the user does not have to manually instrument its
%     code to dump raw data and creates specific post-processing scripts to visualize
%     it 2) PyTracer allows the user to export the raw data traced to extend the
%     analysis. 
% }

% \add{
%     The main challenge addressed by PyTracer is the transparent instrumentation of Python 
%     applications. Indeed, since Python is an interpreted language, one has no complete 
%     view of the function that will be executed by a given program. 
%     Moreover, for Python applications used in HPC environments, one has
%     most of the time no access to source codes or with limited write access.
%     Therefore, PyTracer tackles this issue by instrumenting the function dynamically 
%     thanks to the Python introspection.
% }

% \add{
%     The trace-based approach used by PyTracer permits to limit the 
%     overhead at runtime since the statistics computation is reported to the postprocess analysis.
%     However, it limits the aggregation to deterministic executions,
%     although control-flow analysis can help to merge divergent execution paths.
%     Hence, PyTracer is restricted to single-threaded applications even if its usage
%     is limited in Python due to the Global Interperter Lock (GIL) mechanism
%     that allows only one thread to hold the control of the Python interpreter.
% }

% \subsection{Pytracer workflow}

% \tristan{mention monkey patching}

@@

% \begin{listing}
%     \centering
%     \begin{lstlisting}[language=Python,style=customPython]
% def get_wrapper_function(module, qualname, name, function):
%     """Returns the instrumented function as a string.
%     This string will be compiled with the compile builtin function

%     Parameters:
%         module: Name of the function module
%         qualname: Qualified name of the function
%         name: Name of the function
%         function: Python object function

%     Returns:
%         wrapper_code: The code source of the wrapper
%     """
%     info = (id(function), function_module, qualname)
%     wrapper_code = f"""
%     def {function_wrapper_name}(*args, **kwargs):
%         return generic_wrapper({info},*args,**kwargs)"""
%     return wrapper_code
% \end{lstlisting}
%     \caption{Function to create the instrumented version of a function.
%         \pytracer uses the identifier of the function instead of the 
%         actual function to handle aliases and avoid duplicate instrumentation.}
%     \label{fig:wrapper_creation}
% \end{listing}

% \begin{listing}
%     \centering
%     \begin{lstlisting}[language=Python,style=customPython,]
% def generic_wrapper(self, info, *args, **kwargs):
%     """Generic wrapper saving inputs and outputs of the wrapped function.
%     The id_dict dict keeps a mapping between the original function identifier
%     and itself. Arguments are binded to avoid mispositioning.
%     Returns the actual output of the wrapped function.

%     Parameters:
%         info: Tuple with the function's id, function's module and function's name
%         *args: Positional calling arguments
%         **kwargs: Keyword calling arguments

%     Returns:
%         outputs: Outputs of the wrapped function
%     """
%     fid, module_name, function_name = info
%     function = original_function_cache.id_dict[fid]
%     bind = Binding(function, *args, **kwargs)
%     stack = self.backtrace()
%     time = elements()
%     self.inputs(time, module_name, function_name, function, bind.arguments, stack)
%     outputs = function(*bind.args, **bind.kwargs)
%     self.outputs(time, module_name, function_name, function, outputs, stack)
%     return outputs
% \end{lstlisting}
%     \caption{\pytracer's generic wrapper function that saves
%         the inputs and outputs of the wrapped function.}
%     \label{fig:generic_wrapper}
% \end{listing}

@@

Three  main  approaches  exist  to  detect  numerical  instabilities:
stochastic  arithmetic,  uncertainty  or  sensitivity analysis [23], and random
seed analysis [11].  \add{PyTracer is agnostic to the noise model used and therefore
    works for all methods mentioned}.  For our use cases, we focus on stochastic
arithmetic since it does not make any assumption about the noise model, in
contrast to sensitivity analysis\add{.}

@@

We enabled MCA in Python programs via Verificarlo~\cite{verificarlo}, a
clang-based compiler~\cite{lattner2008llvm} that replaces floating-point
operations by a generic call to a configurable floating-point model. Several
floating-point models are
available~\cite{chatelain2019automatic,chatelain2019outils}, also called
backends. \pytracer leverages ``fuzzy"~\cite{kiar2020comparing}, a collection of
software tools compiled with Verificarlo. In particular, fuzzy provides
MCA-instrumented versions of the Python interpreter as well as the BLAS, LAPACK,
NumPy, SciPy, and scikit-learn libraries, which enables MCA for a wide range of
existing scientific Python programs. Fuzzy is available in Verificarlo's GitHub
organization at
\href{https://github.com/verificarlo/fuzzy}{\url{github.com/verificarlo/fuzzy}}.

@@

We evaluated \pytracer on two widely-used Python libraries:
SciPy~\cite{virtanen2020scipy}, the main library for scientific computing in
Python, and  scikit-learn~\cite{pedregosa2011scikit}, a reference
machine-learning framework that implements a wide range of state-of-the-art
algorithms. For each library, we executed the software examples provided in
their respective GitHub repositories. We traced each example five times, having
activated MCA in the Python interpreter \add{as well as} BLAS, LAPACK, NumPy,
SciPy, and scikit-learn. \add{Instrumenting only the Python interpreter would
    lead to only perturb pure Python floating-point operations and thus miss
    operations in compiled codes (i.e., Cython and shared libraries).} \add{We
    used MCA in Random Rounding (RR) mode, with 5 repetitions}. We used the default virtual
precision in Verificarlo, namely 24 bits for single-precision floating-point
values and 53 bits for double-precision values, which simulates machine error.
We used Python-3.8.5, NumPy-1.19.1, SciPy-1.5.4, and scikit-learn-0.23.2.

% \subsection{Results overview}

% Table~\ref{tab:pytracer_results_summary} summarizes \pytracer's outcome on SciPy
% and scikit-learn examples. Columns \textit{trace} and \textit{parse} report on
% the tracing and aggregation steps with three possible outcomes: 1) green for
% success, meaning that the execution ended well and no issues were raised by the
% application, 2) red for numerical error in the application, meaning that the
% execution raised a Python error before the end, and 3) orange for errors due to
% the fuzzy noise injection that are not considered numerical errors, meaning that
% the execution raised a Python error but due to a perturbation that should not
% have been introduced (i.e. a floating-point value representing an integer).
% Overall, the tracing and parsing of SciPy and scikit-learn examples in Random
% Rounding mode was entirely successful for 32/40 examples and showed an average
% precision loss of 8 bits. The remaining examples failed due to numerical errors
% in the libraries, which we discuss hereafter. The instrumentation with Full MCA
% is more invasive and reduced the number of successful executions to 12/40. Among
% the 28/40 failed executions, one failed due to an actual numerical error
% (\texttt{Bayesian Ridge Regression} example) and the other ones failed due to
% perturbations that should not have been introduced.

% Column \textit{sigbits} in Table~\ref{tab:pytracer_results_summary} summarizes
% the precision for the main outputs of SciPy and scikit-learn examples. We
% measured the precision in significant bits on the final primary example output
% by taking the element-wise mean for non-scalar values and, if the example
% produces many outputs, the average value among all of them. The numerical
% quality across all examples varies from 0 to 52, with an average of 44
% significant bits. For both modes, we note that only two examples failed due to a
% numerical error (\texttt{face\_recognition} for RR and \texttt{Bayesian Ridge
%     Regression} for Full MCA).

@@

Table~\ref{tab:pytracer_results_summary} summarizes \pytracer's outcome on SciPy
and scikit-learn examples. Columns \textit{trace} and \textit{parse} report on
the tracing and aggregation steps with two possible outcomes: 1) green for
success, meaning that the execution ended well and no issues were raised by the
application, and 2) red for numerical error in the application, meaning that the
execution raised a Python error before the end. \remove{, and 3)
    orange for errors due to the fuzzy noise injection that are
    not considered numerical errors, meaning that the execu-
    tion raised a Python error but due to a perturbation that
    should not have been introduced (i.e. a floating-point value
    representing an integer).}
Overall, the tracing and parsing of SciPy and scikit-learn examples in Random
Rounding mode was entirely successful for 32/40 examples\remove{ and showed an average
    precision loss of 8 bits}. The remaining examples failed due to numerical errors
in the libraries, which we discuss hereafter. 
\remove{The instrumentation with Full MCA is
    more invasive and reduced the number of successful exe-
    cutions to 12/40. Among the 28/40 failed executions, one
    failed due to an actual numerical error (Bayesian Ridge
    Regression example) and the other ones failed due to
    perturbations that should not have been introduced.}

@@

% RR: Crash due to FP32 use. Does FP64 converge? No. Does sgsvd converge? Yes

% MCA: arange issue

% File "/usr/local/lib/python3.8/site-packages/sklearn/datasets/\_lfw.py", line
%     169, in \_load\_imgs faces[i, ...] = face ValueError: could not broadcast
%     input array from shape (49,37) into shape (50,37)

% \paragraph{L1 Penalty} This test compares the percentage of zero coefficients
% for logistic regression by varying the penalty (L1, L2 and Elastic-Net) and
% regularization (1, 0.1, 0.01) parameters. A large regularization value gives
% the model more freedom than a small value, which puts more constraints on the
% model.

% of L1, L2 and Elastic-Net penalty.

% Comparison of the sparsity (percentage of zero coefficients) of solutions when
% L1, L2 and Elastic-Net penalty are used for different values of C. We can see
% that large values of C give more freedom to the model. Conversely, smaller
% values of C constrain the model more. In the L1 penalty case, this leads to
% sparser solutions. As expected, the Elastic-Net penalty sparsity is between
% that of L1 and L2. We classify 8x8 images of digits into two classes: 0-4
% against 5-9. The visualization shows coefficients of the models for varying C.

% RR: ok \\
% MCA: stable for L1 and L2, not for Len but coef have less zeroes.

% \paragraph{Lasso and Elastic-Net} This test compares the Lasso and Elastic-Net
% regression models on a corrupted signal.

% Estimates Lasso and Elastic-Net regression models on a manually generated
% sparse signal corrupted with an additive noise. Estimated coefficients are
% compared with the ground-truth.

% RR: Good (48 bits on coefs) \\
% MCA: Good like RR if we exclude one trace that fails due to arange issue

% \paragraph{MNIST classification} This test fits a logistic regression with the
% SAGA solver with the L1 penalty on the MNIST digits classification task
% subset.

% Here we fit a multinomial logistic regression with L1 penalty on a subset of
% the MNIST digits classification task. We use the SAGA algorithm for this
% purpose: this a solver that is fast when the number of samples is
% significantly larger than the number of features and is able to finely
% optimize non-smooth objective functions which is the case with the l1-penalty.
% Test accuracy reaches > 0.8, while weight vectors remains sparse and therefore
% more easily interpretable. Note that this accuracy of this l1-penalized linear
% model is significantly below what can be reached by an l2-penalized linear
% model or a non-linear multi-layer perceptron model on this dataset.

% RR: Very stable \\
% MCA: Parse by excluding a trace that failed due to an arange issue

% \paragraph{Multitask Lasso} This test compares the Lasso and MultiTaskLasso
% regression models on a corrupted signal.

% The multi-task lasso allows to fit multiple regression problems jointly
% enforcing the selected features to be the same across tasks. This example
% simulates sequential measurements, each task is a time instant, and the
% relevant features vary in amplitude over time while being the same. The
% multi-task lasso imposes that features that are selected at one time point are
% select for all time point. This makes feature selection by the Lasso more
% stable.

% RR: Stable (44~53 bits for coef) \\
% MCA: Stable like RR if we exclude a trace that failed due index issue

% \paragraph{Orthogonal matching} This test uses the Orthogonal Matching Pursuit
% algorithm for recovering a sparse signal from a noisy measurement encoded with
% a dictionary.

% RR: Very stable \\
% MCA: Very stable

% \paragraph{PCA decomposition} This test makes a PCA decomposition on a
% synthetic dataset. These figures aid in illustrating how a point cloud can be
% very flat in one direction–which is where PCA comes in to choose a direction
% that is not flat.

% rr: stable (48 bits on components) \\
% mca: stable (48 bits on components)

% \paragraph{Robust Linear Regression} This test compares the linear and  RANSAC
% regression methods on noisy data. In this example we see how to robustly fit a
% linear model to faulty data using the RANSAC algorithm.

% RR: Very stable \\
% MCA: Very stable if we exclude two traces that failed due to arange/index
% issues

% \paragraph{Segmentation Toy} This test uses spectral clustering to separate
% connected circles in an image.

% In this example, an image with connected circles is generated and spectral
% clustering is used to separate the circles.

% In these settings, the Spectral clustering approach solves the problem know as
% ‘normalized graph cuts’: the image is seen as a graph of connected voxels, and
% the spectral clustering algorithm amounts to choosing graph cuts defining
% regions while minimizing the ratio of the gradient along the cut, and the
% volume of the region.

% As the algorithm tries to balance the volume (ie balance the region sizes), if
% we take circles with different sizes, the segmentation fails.

% In addition, as there is no useful information in the intensity of the image,
% or its gradient, we choose to perform the spectral clustering on a graph that
% is only weakly informed by the gradient. This is close to performing a Voronoi
% partition of the graph.

% In addition, we use the mask of the objects to restrict the graph to the
% outline of the objects. In this example, we are interested in separating the
% objects one from the other, and not from the background.

% RR Stable \\
% MCA: Stable if we exclude three traces that failed due to index issue

% Find the optimal separating hyperplane using an SVC for classes that are
% unbalanced.

% We first find the separating plane with a plain SVC and then plot (dashed) the
% separating hyperplane with automatically correction for unbalanced classes.

% RR: stable (48 bits) \\
% MCA: stable (48 bits)

% \paragraph{Tomography} This test compares the LASSO and Ridge regression
% models to reconstruct an image from  projections for a dataset acquired from
% computed tomography.

% This test shows the reconstruction of an image from a set of parallel
% projections, acquired along different angles. Such a dataset is acquired in
% computed tomography (CT).

% Without any prior information on the sample, the number of projections
% required to reconstruct the image is of the order of the linear size l of the
% image (in pixels). For simplicity we consider here a sparse image, where only
% pixels on the boundary of objects have a non-zero value. Such data could
% correspond for example to a cellular material. Note however that most images
% are sparse in a different basis, such as the Haar wavelets. Only l/7
% projections are acquired, therefore it is necessary to use prior information
% available on the sample (its sparsity): this is an example of compressive
% sensing.

% The tomography projection operation is a linear transformation. In addition to
% the data-fidelity term corresponding to a linear regression, we penalize the
% L1 norm of the image to account for its sparsity. The resulting optimization
% problem is called the Lasso. We use the class Lasso, that uses the coordinate
% descent algorithm. Importantly, this implementation is more computationally
% efficient on a sparse matrix, than the projection operator used here.

% The reconstruction with L1 penalization gives a result with zero error (all
% pixels are successfully labeled with 0 or 1), even if noise was added to the
% projections. In comparison, an L2 penalization (Ridge) produces a large number
% of labeling errors for the pixels. Important artifacts are observed on the
% reconstructed image, contrary to the L1 penalization. Note in particular the
% circular artifact separating the pixels in the corners, that have contributed
% to fewer projections than the central disk.

% RR Divergence \\
% MCA Index issue

% Finally, taking into account the points' weight allows for a better separation
% \tristan{is this related to numerical stability? If not, you can ommit the
% comment.}.

% RR Stable (30~48 bits on coef) \\
% MCA Unstable (0 bits on coef)

% \subsection{PyAFQ}

% PyAFQ is a Python package for Automated Fiber Quantification. It aims to find
% delineation of the major fiber tracts in individual human brains and quantify
% the tissue properties within the tracts. PyAFQ \pytracer tracing produces
% traces of 1TB with an average slowdown $\times 6$ compared to the actual
% execution. Moreover, \pytracer was able to visualize the aggregated traces,
% demonstrating its scalability fluently.

@@

\pytracer's tracing overhead consists of (1) module instrumentation time
(initialization overhead) and (2) data write time during the execution (runtime
overhead). The initialization overhead is proportional to the number of
functions and classes traced. For example, we measured an average initialization
time of 8 seconds during our experiments with  NumPy, SciPy, and Scikit-learn.
This overhead can be reduced by excluding intensively used functions such as
\texttt{numpy.array} from the instrumentation.
Figure~\ref{fig:performance_tracing} summarizes \pytracer's tracing overhead for
SciPy and scikit-learn examples. \textit{Slowdown} represents the slowdown of
\pytracer without fuzzy measured as the ratio between the actual execution time
(without \pytracer) and the execution time using \pytracer. The \textit{Slowdown
    RR} \remove{and \textit{Slowdown MCA}} shows the slowdown within the fuzzy environment.
Without fuzzy, \pytracer introduces an average slowdown of $17\times$ for
the examples running less than 8 seconds because \pytracer spends most of its
time in the initialization step. Since this step is a constant factor of the
number of functions instrumented, we also measured the instrumentation cost
itself (\textit{Slowdown (Instrumentation)}) calculated by subtracting the
initialization time to \textit{Slowdown}. We then obtain an average slowdown of
$3.5\times$. Finally, the \remove{fuzzy instrumentation} \add{using MCA
    operations (through fuzzy) instead of the native ones} increases the average
slowdown to $37\times$ for RR\remove{and $43\times$ for Full MCA}.

@@

\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{figure/performance.pdf}
    \caption{
        \pytracer tracing overheads. \textit{Slowdown} represents the slowdown
        of \pytracer without fuzzy measured as the ratio between the actual
        execution time (without \pytracer) and the execution time using
        \pytracer. \textit{Slowdown (Instrumentation)} show the actual
        instrumentation cost computed by subtracting the initialization step.
        The \textit{Slowdown (RR)} \remove{and \textit{Slowdown (MCA)} are} \add{is} the slowdown
        of \pytracer using the fuzzy environment with \add{the} RR \remove{and Full MCA} mode\remove{s}. The
        \remove{first} \add{last} four examples measure the initialization
        steps depending on the modules instrumented. 
        % \tristan{there's too much text in this caption, move some of it to the text, the caption should just explain the difference between the 4 slowdowns. From the text I'm not sure what is the difference between slowdown and slowdown(RR)/slowdown(MCA).}
    }
    \label{fig:performance_tracing}
\end{figure}

@@

Through our analysis of SciPy and scikit-learn, we demonstrated \pytracer's
scalability, automaticity, and applicability to a wide range of Python program.
The relatively low instrumentation overhead of 3.5$\times$ on average enables
the analysis of real-scale applications. \pytracer's instrumentation is fully
automatic and does not require manual intervention from the user. Furthermore,
\pytracer works on any Python libraries, with any data format supported by
pickle and any numerical noise model.

@@

\pytracer allows users to pinpoint the root causes of numerical instabilities by
working at the function level. Numerical stability evaluations are complex since
calculations are strongly interdependent. Function-level instrumentation limits
the size of the traces since only inputs and outputs of a function are dumped.
Moreover, it decreases instrumentation overhead and helps the user identify
instabilities quickly. Working at the function level is thus a suitable
compromise between analysis conducted at the floating-point operation level and
user debug prints. Indeed,  fine-grained analyses are costly and hard to
interpret as they overwhelm users with information. In contrast, looking at
intermediate results through print statements is lightweight but not reliable
since developers may not systematically instrument unstable functions. Finally,
the SciPy environment wraps many computation kernels written in C, such as BLAS
or LAPACK. By working at the function level, \pytracer allows the user to
evaluate the numerical precision of such kernels and their impact on the code
without getting deep into technical implementation aspects.

@@

\pytracer can also be useful for selecting hyperparameters, as is the case in
machine learning. Indeed, it is common for data scientists to explore a range of
hyperparameters and select the ones giving the best scores. With \pytracer, the
user has a practical way to quantify numerical quality and understand why
hyperparameters produce inconsistent results instead of tweaking hyperparameters
to make solvers converge. For instance, the Face Recognition example
(Figure~\ref{fig:face_recognition_svd}) shows how numerical precision could
inform the selection of the number of components in dimensionality reduction.

The visualization provides a global view of the code and interactions between
computations. In addition, the Plotly dashboard helps navigate through the code
easily, without the need for ad-hoc scripts to visualize data. Finally, the
dashboard can be extended to support other data formats to visualize. At this
stage, \pytracer currently supports n-dimensional NumPy arrays as well as native
Python data.

@@

Finally, the wide range of results precision obtained from our experiments
demonstrate that looking at the numerical variability is essential and that such
analyses would benefit from being conducted systematically. We showed that
well-known scientific libraries are prone to numerical error even though they
have been well tested for decades. The omnipresence of these libraries in
scientific codes should thus impel users to integrate numerical analysis in
their development workflow in addition to testing their implementation.

@@

% \label{sec:detecting-instabilities} \paragraph{Dynamically detecting numerical
% instabilities} \label{sec:mca}
% :
%  \Yohan{Either detail this section with UA/SA or remove it since MCA is
%  introduced in 'noise injection' subsection. } \tristan{I agree, let's just
%  remove it, the core of the contribution is on the tracing, not in the
%  detection of instabilities.}

>>>>

% \ifANNOTATED
\definecolor{ao(english)}{rgb}{0.0, 0.5, 0.0}
\DeclareRobustCommand{\add}[1]{\textcolor{ao(english)}{#1}}%\sethlcolor{lightgreen}\hl{#1}}
\DeclareRobustCommand{\remove}[1]{\textcolor{lightred}{\sout{#1}}}
\DeclareRobustCommand{\addCite}[1]{\textcolor{lightgreen}{\mbox{#1}}}
\newcommand{\TG}[1]{\color{red}\textsc{From Tristan}: #1\xspace\color{black}}
% \else
% \DeclareRobustCommand{\add}[1]{#1}
% \DeclareRobustCommand{\remove}[1]{}
% \DeclareRobustCommand{\addCite}[1]{#1}
% \newcommand{\TG}[1]{}
% \fi

@@

\pytracer estimates the variability of floating-point variables by
combining program traces obtained with different numerical perturbations,
which requires (\add{a}\remove{1}) generating numerically-perturbed
executions, (\add{b}\remove{2}) tracing floating-point computations along
program executions, and (\add{c}\remove{3}) visualizing stability
evaluations. \add{PyTracer addresses (a) by relying on
Fuzzy~\addCite{\cite{kiar2020comparing}}, a pre-existing collection of scientific
computing libraries instrumented with Monte-Carlo arithmetic through the
Verificarlo tool~\cite{verificarlo}, (b) by dynamically instrumenting
Python functions, modules, and classes through meta-programming, and (c) by
visualizing stability measures throughout program executions with an
interactive Plotly dashboard. \mbox{\pytracer} is available at
\mbox{\href{https://github.com/big-data-lab-team/pytracer.git}{github.com/big-data-lab-team/pytracer.git}}
under MIT License.} \remove{\mbox{\pytracer} addresses these requirements
with (1) a Fuzzy Python interpreter instrumented with Monte-Carlo
arithmetic, (2) a dynamic instrumentation of Python functions, modules, and
classes through meta-programming, (3) an interactive Plotly dashboard to
visualize stability measures throughout program executions. }

@@

\add{
    To detect numerical instabilities, \pytracer works in three main steps:
\begin{enumerate}
\item It instruments Python programs and their module dependencies
automatically, to save the input and output values of functions in a trace
file.
\item It executes the instrumented program several times using Monte-Carlo
Arithmetic to inject perturbations in function outputs.
\item It aggregates all trace files to compute statistics about the traced
values, and visualize them in a GUI.
\end{enumerate}
\pytracer adopts the workflow shown in Figure~\ref{fig:workflow}.
At step 1, \pytracer uses monkey patching to replace the functions called
by the application with a wrapper that saves their arguments and returned
values in a trace file. At step 2, it executes the application multiple
times with Monte-Carlo Arithmetic as implemented in the
Fuzzy~\cite{kiar2020comparing} environment. It then
aggregates the resulting trace files to compute summary statistics on
traced variable. At step 3, a Plotly dashboard provides interactive
visualizations highlighting unstable code sections.
}

@@

@@

Three  main  approaches  exist  to  detect  numerical  instabilities:
stochastic  arithmetic,  uncertainty  or  sensitivity analysis [23], and random
seed analysis [11].  \add{PyTracer is agnostic to the noise model used and can therefore
be used for all these methods}.  For our use cases, we focus on stochastic
arithmetic since it does not make any assumption about the noise model, in
contrast to sensitivity analysis\add{.}

@@

We enabled MCA in Python programs via Verificarlo~\cite{verificarlo}, a
clang-based compiler~\cite{lattner2008llvm} that replaces floating-point
operations by a generic call to a configurable floating-point model. Several
floating-point models are
available~\cite{chatelain2019automatic,chatelain2019outils}, also called
backends. \pytracer leverages Fuzzy~\cite{kiar2020comparing}, a collection of
software tools compiled with Verificarlo. In particular, Fuzzy provides
MCA-instrumented versions of the Python interpreter as well as the BLAS, LAPACK,
NumPy, SciPy, and scikit-learn libraries, which enables MCA for a wide range of
existing scientific Python programs. Fuzzy is available in Verificarlo's GitHub
organization at
\href{https://github.com/verificarlo/fuzzy}{\url{github.com/verificarlo/fuzzy}}.

@@

We evaluated \pytracer on two widely-used Python libraries:
SciPy~\cite{virtanen2020scipy}, the main library for scientific computing in
Python, and  scikit-learn~\cite{pedregosa2011scikit}, a reference
machine-learning framework that implements a wide range of state-of-the-art
algorithms. For each library, we executed the software examples provided in
their respective GitHub repositories. We traced each example five times, having
activated MCA through Fuzzy in the Python interpreter \add{as well as} BLAS, LAPACK, NumPy,
SciPy, and scikit-learn. \add{These libraries are written in large parts in a 
compiled language other than Python (such as C or Cython) and their execution is not 
perturbed by the sole use of Fuzzy Python. We chose to activate MCA in these libraries 
to provide an overall picture of numerical uncertainty in the tested examples. 
Depending on the use-case of interest, one might choose to use MCA only in 
a subset of these libraries or even just in the Python interpreter.} \add{We
    used MCA in Random Rounding (RR) mode}. We used the default virtual
precision in Verificarlo, namely 24 bits for single-precision floating-point
values and 53 bits for double-precision values, which simulates machine error.
We used Python-3.8.5, NumPy-1.19.1, SciPy-1.5.4, and scikit-learn-0.23.2.

@@

Table~\ref{tab:pytracer_results_summary} summarizes \pytracer's outcome on SciPy
and scikit-learn examples. Columns \textit{trace} and \textit{parse} report on
the tracing and aggregation steps with two possible outcomes: 1) green for
success, meaning that the execution ended well and no issues were raised by the
application, and 2) red for numerical error in the application, meaning that the
execution raised a Python error before the end. \remove{, and 3)
    orange for errors due to the Fuzzy noise injection that are
    not considered numerical errors, meaning that the execu-
    tion raised a Python error but due to a perturbation that
    should not have been introduced (i.e. a floating-point value
    representing an integer).}
Overall, the tracing and parsing of SciPy and scikit-learn examples in Random
Rounding mode was entirely successful for 32/40 examples\remove{ and showed an average
    precision loss of 8 bits}. The remaining examples failed due to numerical errors
in the libraries, which we discuss hereafter. 
\remove{The instrumentation with Full MCA is
    more invasive and reduced the number of successful exe-
    cutions to 12/40. Among the 28/40 failed executions, one
    failed due to an actual numerical error (Bayesian Ridge
    Regression example) and the other ones failed due to
    perturbations that should not have been introduced.}

@@

@@

\pytracer's tracing overhead consists of (1) module instrumentation time
(initialization overhead) and (2) data write time during the execution (runtime
overhead). The initialization overhead is proportional to the number of
functions and classes traced. For example, we measured an average initialization
time of 8 seconds during our experiments with  NumPy, SciPy, and Scikit-learn.
This overhead can be reduced by excluding intensively used functions such as
\texttt{numpy.array} from the instrumentation.
Figure~\ref{fig:performance_tracing} summarizes \pytracer's tracing overhead for
SciPy and scikit-learn examples. \textit{Slowdown} represents the slowdown of
\pytracer without fuzzy measured as the ratio between the actual execution time
(without \pytracer) and the execution time using \pytracer. The \textit{Slowdown
    RR} \remove{and \textit{Slowdown MCA}} shows the slowdown within the Fuzzy environment.
Without Fuzzy, \pytracer introduces an average slowdown of $17\times$ for
the examples running less than 8 seconds because \pytracer spends most of its
time in the initialization step. Since this step is a constant factor of the
number of functions instrumented, we also measured the instrumentation cost
itself (\textit{Slowdown (Instrumentation)}) calculated by subtracting the
initialization time to \textit{Slowdown}. We then obtain an average slowdown of
$3.5\times$. Finally, the \remove{Fuzzy instrumentation} \add{using MCA
    operations (through Fuzzy) instead of the native ones} increases the average
slowdown to $37\times$ for RR\remove{and $43\times$ for Full MCA}.

@@

\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{figure/performance.pdf}
    \caption{
        \pytracer tracing overheads. \textit{Slowdown} represents the slowdown
        of \pytracer without Fuzzy measured as the ratio between the actual
        execution time (without \pytracer) and the execution time using
        \pytracer. \textit{Slowdown (Instrumentation)} show the actual
        instrumentation cost computed by subtracting the initialization step.
        The \textit{Slowdown (RR)} \remove{and \textit{Slowdown (MCA)} are} \add{is} the slowdown
        of \pytracer using the Fuzzy environment with \add{the} RR \remove{and Full MCA} mode\remove{s}. The
        \remove{first} \add{last} four examples measure the initialization
        steps depending on the modules instrumented. 
        % \tristan{there's too much text in this caption, move some of it to the text, the caption should just explain the difference between the 4 slowdowns. From the text I'm not sure what is the difference between slowdown and slowdown(RR)/slowdown(MCA).}
    }
    \label{fig:performance_tracing}
\end{figure}

@@

Through our analysis of SciPy and scikit-learn, we demonstrated \pytracer's
scalability, automaticity, and applicability to a wide range of Python programs.
The relatively low instrumentation overhead of 3.5$\times$ on average enables
the analysis of real-scale applications. \pytracer's instrumentation is fully
automatic and does not require manual intervention from the user. Furthermore,
\pytracer works with any Python libraries, with any data format supported by
pickle and any numerical noise model.

@@

\pytracer allows users to pinpoint the root causes of numerical instabilities by
working at the level of Python functions. Numerical stability evaluations are complex since
calculations are strongly interdependent. Function-level instrumentation limits
the size of the traces since only inputs and outputs of a function are dumped.
Moreover, it decreases instrumentation overhead and helps the user identify
instabilities quickly. Working at the function level is thus a suitable
compromise between analyses conducted at the floating-point operation level and
user debug prints. Indeed, operation-level analyses are costly and hard to
interpret as they overwhelm users with information. In contrast, looking at
intermediate results through print statements is lightweight but not reliable
since developers may not systematically instrument unstable functions.

\add{\pytracer can be used with different types of numerical perturbations,
including random seeds, data noise, or stochastic arithmetic. The
use of Verificarlo to implement stochastic arithmetic through the Fuzzy
environment enables to study the impact of numerical perturbations
introduced in non-Python dependencies while preserving the high-level view
of Python variables and functions. For instance, the SciPy environment
wraps many computation kernels written in C, such as BLAS or LAPACK, which
have a considerable impact on the stability of applications using them.}

\add{
Enabling Monte-Carlo arithmetic in such non-Python dependencies requires
their recompilation with Verificarlo, which usually requires human
intervention. The Fuzzy environment provides pre-instrumented version of
NumPy, BLAS, LAPACK, GNU libmath, SciPy, and Scikit-learn, which, given the
prevalence of these libraries in scientific Python codes, should cover a
large ground of non-Python dependencies encountered in scientific Python programs.
}

@@

\pytracer might also be useful for selecting hyperparameters, as is the case in
machine learning. Indeed, it is common for data scientists to explore a range of
hyperparameters and select the ones giving the best scores. With \pytracer, the
user has a practical way to quantify numerical quality and understand why
hyperparameters produce inconsistent results instead of tweaking hyperparameters
to make solvers converge. For instance, the Face Recognition example
(Figure~\ref{fig:face_recognition_svd}) shows how numerical precision could
inform the selection of the number of components in dimensionality reduction.

\pytracer's visualization provides a global view of the code and interactions between
computations. In addition, the Plotly dashboard helps navigate through the code
easily, without the need for ad-hoc scripts to visualize data. The
dashboard can be extended to support other data formats to visualize. At this
stage, \pytracer currently supports n-dimensional NumPy arrays as well as native
Python data.

@@

Finally, the wide range of results precision observed in our experiments
suggests that such
analyses would benefit from being conducted systematically. We showed that
well-known scientific libraries are prone to numerical errors even though they
have been well tested for decades. The prevalence of these libraries in
scientific codes should thus impel users to integrate numerical analysis in
their development workflow in addition to testing their implementation.

@@
